\section{Evaluation and Implementation}
\label{sec:imp-eval}


We developed an automated chain of translation and interpretation of C program to evaluate our semantics. We applied OSL to 20 real-world open source snippets programs, developed by external developers. All these programs have different sizes and complexities. \autoref{tab:evaluation-table} presents size and timing information of some evaluated programs. These benchmarks had been made with a $Intel(R) Core(TM)$ $i7-7700HQCPU @ 2.80GHz$, $16GB$ of RAM and 4 physical cores.

\begin{table}[h]
  \centering
  \begin{tabular}{ | c | c | c | c | c | c | }
  \hline
    Program & Use std-libs & Lines of code  & Transpiled size & Exec time & CPU Usage \\ \hhline{|=|=|=|=|=|=|}
    ASCII Chess Game & yes  & 566 & 532 & 8s & 32\% \\ \hline
    Simple RSA & yes & 230 & 261 & 4s & 25\% \\ \hline
c    Echo HTTP server & yes &  101 & 95 & 3s & 23\% \\ \hline
    8-queens solver &  no & 119 & 135 & 3s & 26\% \\
    \hline 
  \end{tabular}
  \vspace{\baselineskip}
  \caption{Outline evaluation of OSL}
  \label{tab:evaluation-table}
\end{table}

In \autoref{tab:evaluation-table}, the size of programs does not include the code add by the inclusion of C standard library. Moreover,
\oslt~apply the GCC $preprocessor$ on program input before translating them. Currently, OSL does not provide complete
support of the C standard library. The complexity of translating C program with \oslt~lies in the complexity that a C expression can have.
As we do not alter and annotate C input programs, some C expressions require deeper statistical analysis and a greater covering of the standard library than the one we proposed currently.\\

The semantics of \oslos~has been implemented in version 3.5 of $\mathbb{K}$-Framework.
The compilation of the \oslos~model by $\mathbb{K}$-Framework produces a graph of 40 states and 1247 transitions.
The translator \oslt, however, is a Rust binary program.
The latest version 5.0 of $\mathbb{K}$-Framework significantly speeds up the executions of the executable semantics and compilation time.
But the latest version is a breaking change version. In the short-term, we plan to migrate to the latest version to improve the executable semantics execution time.
The code source is publicly available \cite{OSL}.